---
title: "Class19"
author: "Evelyn Diaz - PID:A15576617"
date: "11/30/2021"
output: html_document
---

#Section 1, Proportion of G/G in a population

Downloaded a CSV file from Ensemble<
https://uswest.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39827525-39962692;v=rs8067378;vdb=variation;vf=105535077#373531_tablePanel>

here we read this csv file

```{r}
mxl<- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)

```

```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.)/ nrow(mxl) *100
```

Now let's look at a different population, I picked the GBR.

```{r}
gbr<- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(gbr)
```

Find proportion of G|G

```{r}
round(table(gbr$Genotype..forward.strand.) / nrow(gbr) *100, 2)
```

This variant that is associated with childhood asthma is more frequent in the GBR population than the MXL population.

#Homework, section 4
One sample is obviously not enough to know what is happening in a population. You are
interested in assessing genetic differences on a population scale.

13)How many samples do we have?

```{r}
expr<- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```


```{r}
nrow(expr)
```

```{r}
table(expr$geno)
```



```{r}
library(ggplot2)
```

14) Let's make a boxplot:
```{r}
ggplot(expr) + aes(x= geno, y=exp, fill=geno) +
  geom_boxplot(notch=T)

```


















