---
title: "Class 12 Structural Bioninformatics II"
author: "Evelyn Diaz - PID:A15576617"
date: "11/4/2021"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

\#Comparative analysis of protein structures

using the bio3d package.

```{r}
library(bio3d)

pdb<- read.pdb("1hel")
pdb
```

Let's use a bioinformaticcs method called NMA (Normal Mode Analysis) to predict the dynamics (flexibility) of the enzyme.

```{r}
modes<- nma(pdb)
plot(modes)
```

Make a "move" of its predicted motion. We often call this a "trajectory".

![](vmdscene2.bmp)

\#Analysis of ADK

```{r}
aa<- get.seq("1ake_A")
aa
```

```{r}
#Run Blast
#blast <- blast.pdb(aa)

#Save all this stuff
# save(blast, files, file="mydata.RData")
load("mydata.RData")
```

```{r}
hits<- plot(blast)
```

```{r}
hits$pdb.id
```

```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','4X8M_A','6S36_A','6RZE_A','4X8H_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','4NP6_A','3GMT_A','4PZL_A')

```

```{r}
# Download releated PDB files
#files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

Multiple structure alignment

```{r}
pdbs <- pdbaln(files, fit = TRUE)
```

```{r}
pdbs

```

```{r}
# Vector containing PDB codes for figure axis
ids <- basename.pdb(pdbs$id)

# Draw schematic alignment
plot(pdbs, labels=ids)
```

\#PCA

We will use the bio3d PCA function which is designed for protein structure data.

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```

Make a trajectory visualization of the motion captured by nthe first principal component


![](images/PC1.bmp)


![](images/Actin.bmp)\# Visualize first principal component
pc1 \<- mktrj(pc.xray, pc=1, file="pc_1.pdb")
pc1
