---
title: "Class 6 R Functions"
author: "Evelyn Diaz - PID:A15576617"
date: "10/14/2021"
output: pdf_document
---

## A play with Rmarkdown
This is some plain text. I can make things **bold**. I can also make *things italic*.

```{r}
#This is a code chunk
plot(1:10)

```


## R Functions
In today's class, we are going to write a function together that grades some student work. 

Questions for today:

> **Q1**. Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an NA value) this can be used as a score to be potentially dropped. Your final function should be adquately explained with code comments and be able to work on an example class gradebook such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]



```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```


Let's start with student1 and fin their average score.
```{r}
mean(student1)
```

But we want to drop the lowest score...
We could try the **min()** function.

```{r}
min(student1)
```

The **which.min()** function looks useful:

```{r}
which.min(student1)
```
This gives the position of the lowest score

```{r}
#This would be the lowest score
student1[which.min(student1)]
```
To drop this value I can use minus

```{r}
student1[-which.min(student1)]
```

Let's now use mean() to get the avergae mines the lowest score.

```{r}
mean(student1[-which.min(student1)])
```

It gives NA if student is missing hw.
```{r}
mean(student2[-which.min(student2)])
```

We need to remove the NA elements of the vector
```{r}
#which.min(student2)
mean(student2[-which.min(student2)], na.rm=TRUE)
```

This is not what we want, it dropped the 80 (i.e. the lowest number and not the NA which is the missing homework).

Let's look at student3
```{r}
mean(student3[-which.min(student3)], na.rm=TRUE)
```

One approach is to replace the NA with zero. 

Let's try with student2
```{r}
is.na(student2)
```

The **is.na()** function returns a logical vector where TRUE elements represent where the NA values are. 

```{r}
which(is.na(student2))
```
Now let's make the NA values into zeros.

```{r}
student.prime<- student2
student.prime
student.prime[which(is.na(student.prime))]= 0
student.prime
```


Now we put it all together to get the average score dropping hte lowest where we map NA valued to zero. 
```{r}
student.prime<- student2
student.prime
student.prime[which(is.na(student.prime))]= 0
student.prime
mean(student.prime[-which.min(student.prime)])
```

```{r}
student.prime
```

```{r}
mean(c(100, 90, 90, 90, 90, 97, 80))
```

It works! Check student3

```{r}
student.prime<- student3
student.prime
student.prime[which(is.na(student.prime))]= 0
student.prime
mean(student.prime[-which.min(student.prime)])
```

We got our working snippet! Let's simplify:

```{r}
x<- student3
#Map NA values to zero
x[which(is.na(x))]= 0
#Find the mean wihtout the lowest value
mean(x[-which.min(x)])

```

Now we can use this as the body of my function.
```{r}
grade<- function(x) {
  #Make sure our scores are all numbers
  x<- as.numeric(x)
  
#Map NA values to zero
  x[which(is.na(x))]= 0
#Find the mean wihtout the lowest value
  mean(x[-which.min(x)])}
```

```{r}
grade(student1)
```

```{r}
grade(student3)
```

```{r}
grade(student2)
```

Now we read the full gradebook CSV file
```{r}
scores<- read.csv("https://tinyurl.com/gradeinput", row.names=1)
scores

```

```{r}
grade(scores[10,])


```

```{r}
is.numeric(student1)
```

```{r}
is.numeric(scores[10,])
```

```{r}
grade(as.numeric(scores[2,]))
```

```{r}
as.numeric(c(1,2,NA,4,5))
```

Now grade all students by using **apply()** function.
```{r}
ans<- apply(scores,1, grade)
ans
```

# Q2 A: Who is the top scoring student overall in the gradebook?
```{r}
which.max(ans)
```

The top scoring student overall in the gradebook is Student-18!


# Q3:From your analysis of the gradebook, which homework was toughest on students (i.e. ob- tained the lowest scores overall?
```{r}
apply(scores, 2, mean, na.rm=TRUE)
```
A: Homework three was the toughest on students. 



```{r}
library(ggplot2)
```






