---
title: "Class 16"
author: "Evelyn Diaz - PID:A15576617"
date: "11/23/2021"
output: html_document
---

#Background
As we approach a period of trvael and larger gatherings, lets have a loot at vaccination rates across the state. 

We will take data from the CA.gov site here: 

We will start by downloading the most recently dated “Statewide COVID-19 Vaccines Administered by ZIP Code” CSV file from: https://data.ca.gov/dataset/covid-19-vaccine-progress-dashboard-data-by-zip-code

```{r}
# Import vaccination data
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```

##Quick look at the data structure 

As before, we can use the **skim** function to quickyl overview of the dataset.


```{r}
library(skimr)
```


```{r}
skimr::skim(vax)
```



## Ensure the date column is useful

We will use the **lubridate** package, which can make life allot easier when dealing with dates and times

```{r}
#install.packages("lubridate")
```

```{r}
library(lubridate)
```

```{r}
today()
```

Here we make out `as_of_date` column lubridate format...

```{r}
# Speciffy that we are using the Year-mont-day format
vax$as_of_date <- ymd(vax$as_of_date)

```

Now I can do useful math with dates more easily:

> Q. How many days since the first entry?

```{r}
today() - vax$as_of_date[1]
```

> Q. How many days since the last entry?

```{r}
today() - vax$as_of_date[nrow(vax)]
```

> Q9. How many days between the first and last entry in the dataset?

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

>. Q10 How many unique dates are in the dataset (ie how many different dates are detailed)?

```{r}
length(unique(vax$as_of_date))
```

This sounds good

```{r}
46*7
```

##WOrking with zipcodes 
We will iuse **zipcodeR** package to help make snese of zip codes. 



```{r}
#install.packages("zipcodeR")
```

for example

```{r}
library(zipcodeR)
```

```{r}
geocode_zip('92037')
```


Calculate the distance between the centroids of any two ZIP codes in miles, e.g.

```{r}
zip_distance('92037','92109')
```


More usefully, we can pull census data about ZIP code areas (including median household income etc.). For example:
```{r}
reverse_zipcode(c('92037', "92109") )

```


```{r}
# Pull data for all ZIP codes in the dataset
zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
head(zipdata)
```


##Focus on San Diego County

```{r}
table(vax$county)
```

We will subset with base R
```{r}
inds<- vax$county == "San Diego"

head(vax[inds,])
```

But let's use **dplyr** package and its **filter()** function:

```{r}
library(dplyr)
```

```{r}
sd<- filter(vax, county == "San Diego")
head(sd)
```

How many entries are there for San Diego County?
```{r}
nrow(sd)
```


```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
head(sd.10)
```

> Q11. How many distinct zip codes are listed for San Diego County? 

```{r}
length(unique(sd$zip_code_tabulation_area))
```

>Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

```{r}
ind<- which.max(sd$age12_plus_population)
sd[ind,]
```

What is the population in the 92037 ZIP code area?

```{r}
filter(sd, zip_code_tabulation_area == "92037") [1,]
```

> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”?

```{r}
sd.now<- filter(sd, as_of_date == "2021-11-09")
mean(sd.now$percent_of_population_fully_vaccinated, na.rm=T)
```

We can look at the 6 number summary:
```{r}
summary(sd.now$percent_of_population_fully_vaccinated, na.rm=T)
```

```{r}
library(ggplot2)
```


>Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?

Using base R  plots:

```{r}
hist(sd.now$percent_of_population_fully_vaccinated, na.rm=T)
```

Using ggplot

```{r}
ggplot(sd.now)+
  aes(percent_of_population_fully_vaccinated)+
  geom_histogram(bins=15)
```

##Focus on UCSD/La Jolla

What about 92037 - UCSD/ La Jolla?
```{r}
filter(sd.now, zip_code_tabulation_area == "92037")

```

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

>Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
ggplot(ucsd) +
  aes(x= as_of_date,
      y= percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x= "Date", y="Percent Vaccinated")
```

>Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
#mean(vax.36$percent_of_population_fully_vaccinated)
```


```{r}
ggplot(ucsd) +
  aes(x= as_of_date,
      y= percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  geom_hline(yintercept = 0.66) +
  ylim(c(0,1)) +
  labs(x= "Date", y="Percent Vaccinated")
```

>Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”?

```{r} 
#summary(vax.36.all)
```

>Q18. Using ggplot generate a histogram of this data.

```{r}
#hist(vax.36$percent_of_population_fully_vaccinated)
```



##Comparing 92037 to other similar sized areas?

Population in the 92037 ZIP code area
```{r}
ucsd[1,]$age5_plus_population
```

First we need to subset the full `vax` dataset to include only ZIP code areas with a popualtion as large as 92037.

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36.all <- filter(vax, age5_plus_population > 36144)
nrow(vax.36.all)
                

```


how many other zip code areas
```{r}
length(unique(vax.36.all$zip_code_tabulation_area))
```

```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```



>Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.


Let's make a final figure that shows all these zipcodes. 

```{r}
ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  geom_hline(yintercept = 0.66, color= "red")
```





