---
title: "Class 9"
author: "Evelyn Diaz - PID:A15576617"
date: "10/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r}


# Save your input data file into your Project directory
fna.data <- "WisconsinCancer.csv"

# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv("WisconsinCancer.csv", row.names=1)

head(wisc.df)
```


```{r}
# We can use -1 here to remove the first column
wisc.data <- wisc.df[,-1]
 head(wisc.data)
```


```{r}
# Create diagnosis vector for later 
diagnosis <- as.factor(c(wisc.df[,'diagnosis'])) 
diagnosis
dim(wisc.data)
is.factor(diagnosis)
```

#Q1. How many observations are in this dataset?

## There are 569 observations in this dataset.
```{r}
str(wisc.data)
```

#Q2. How many of the observations have a malignant diagnosis?
##There are 212 observations that have a malignant diagnosis. 
```{r}
table(diagnosis)
```

#Q3. How many variables/features in the data are suffixed with _mean?
##There are 10 variables in the data that are suffixed with "_mean". 

```{r}

colnames<- c(colnames(wisc.data))


length(grep("_mean", colnames))
```

```{r}
# Check column means and standard deviations
colMeans(wisc.data)

apply(wisc.data,2,sd)

```

```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale= T)
summary(wisc.pr)
```

#Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?
 
##From the results, 44.27% of the variance is captured by the first principal component (PC1).

#Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
#Three PCs are required to describe at least 70% of the orginal population. 

#Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
#Seven PCs are required to describe at least 90% pf the original population. 

#Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?
## What stands out to me about this grpah, is that instead of plots, i see a lot of labels, it is very difficult to understand because everything is messy and compacted. I think this is because the labels were used as plots, but I think this could be better if the plots were easier to see. I think this grpah is trying to show too much information at once.  

```{r}
biplot(wisc.pr)
```

```{r}
# Scatter plot observations by components 1 and 2
plot(wisc.pr$x, col = diagnosis , 
     xlab = "PC1", ylab = "PC2")
```


#Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots?
## I notice that there are plots, that are on the grpah in a mcuh more clean way; they are easier to read and they are sorted by color either being malignant or benign. 

```{r}
# Repeat for components 1 and 3
plot(wisc.pr$x[,1], wisc.pr$x[,3], col = diagnosis, 
     xlab = "PC1", ylab = "PC3")

```


```{r}
# Create a data.frame for ggplot
df <- as.data.frame(wisc.pr$x)
df$diagnosis <- diagnosis

# Load the ggplot2 package
library(ggplot2)

# Make a scatter plot colored by diagnosis
ggplot(df) + 
  aes(PC1, PC2, color= diagnosis) + 
  geom_point()
```


#Section 3
```{r}
# Calculate variance of each component
pr.var <- wisc.pr$sdev^2
head(pr.var)
## [1] 13.281608  5.691355  2.817949  1.980640  1.648731  1.207357
```

```{r}
# Variance explained by each principal component: pve
pve <- pr.var / sum(pr.var)

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```

```{r}
## ggplot based graph

library(factoextra)
fviz_eig(wisc.pr, addlabels = TRUE)
```


#Q9. For the first principal component, what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature concave.points_mean?
##The component of the loading vector for concave.points_mean is -0.2608538.  
```{r}
wisc.pr$rotation["concave.points_mean", 1]
```

#Q10. What is the minimum number of principal components required to explain 80% of the variance of the data?
## The minimum number of principal components requwired to explain 80% of the variance of the data is 4. 
```{r}
var<- summary(wisc.pr)
sum(var$importance[3,] < 0.8)
```





